Weak interspecific interactions in a sagebrush steppe? Conflicting evidence from observations and experiments

### 2018-03-22 removals/manuscript/removal_ms.tex
First pass at final version

\title{Weak interspecific interactions in a sagebrush steppe\new{? Conflicting evidence from observations and experiments} }

@@

Stable coexistence requires intraspecific limitations to be stronger than interspecific limitations. The greater the difference between intra- and interspecific limitations, the more stable the coexistence, and the weaker the competitive release any species should experience following removal of competitors. We conducted a removal experiment to test whether a previously estimated model, showing surprisingly weak interspecific competition for four dominant species in a sagebrush steppe, accurately predicts competitive release. Our treatments were 1) removal of all perennial grasses and 2) removal of the dominant shrub, \textit{Artemisia tripartita}. We regressed survival, growth and recruitment on the locations, sizes, and species identities of neighboring plants, along with an indicator variable for removal treatment. If our ``baseline" regression model, which accounts for local plant-plant interactions, accurately explains the observed responses to removals, then the removal coefficient should be non-significant. \new{For survival, the removal coefficients were never significantly different from zero, and only \textit{A. tripartita} showed a (negative) response to removals at the recruitment stage. For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated interspecific competition. For all three grass species, population models based on the vital rate regressions that included removal effects projected 1.4 to 3-fold increases in equilibrium population size relative to the baseline model (no removal effects).} 
However, we found no evidence of higher response to removal in quadrats with higher pretreatment cover of \textit{A. tripartita}, or by plants experiencing higher pre-treatment crowding by \textit{A. tripartita}, 
raising questions about the mechanisms driving the positive response to removal.
\new{While the experimental results showed greater competitive release than our baseline models predicted, we will need more tightly controlled experiments focused on underlying mechanisms to conclusively validate or reject our model's predictions.}

%The ability of models based on observational data to predict population responses to experimental removals increased our confidence that interspecific interactions are in fact weak for at least three of our four study species. The implication is that factors besides interspecific feedbacks play important roles in determining the structure and dynamics of this community. The success of the models also supports the use of multispecies models based on long-term observational data for characterizing competitive interactions.

\textbf{\large{Keywords:}} Coexistence, competition, integral projection model, removal experiment, sagebrush steppe.

@@

In spring 2011, in a large exclosure containing six of the historical permanent quadrats, we established an additional 16 quadrats for the removal treatments. New quadrats were selected to be similar to historical plots, but also needed to have enough of the focal species present so that removal experiments could be conducted. We evaluated potential locations for the new quadrats near the original quadrats and then used only those locations with at least 5\% canopy cover of \textit{Artemisia tripartita} and at least 3\% combined basal cover of the three focal perennial grasses. We rejected locations falling on hill slopes, areas with greater than 20\% bare rock, or with over 10\% cover of the woody shrubs \textit{Purshia tridentata} or \textit{Amelanchier utahensis}. We mapped the new quadrats in June, 2011 and then implemented the removal treatments in September, 2011. We randomly assigned 8 of the sixteen new quadrats to the sagebrush removal treatment, and the remaining 8 to the grass removal treatment. We cut sagebrush stems at ground level and applied a 5\% solution of Roundup Pro (Glyphosate, Monsanto) to the cut stems. In the grass removal plots, we painted the same herbicide solution on all perennial grasses. We removed plants 50 cm beyond the quadrat boundaries to minimize edge effects. The sagebrush removal was virtually 100\% effective; grass removals required additional herbicide applications in April and May of each year, with fewer grasses remaining each year. Appendix Fig. \ref{fig:photos} shows photographs of example quadrats before and after treatment. We repeated mapped censuses of all plots each June from 2012-2016.

@@

Note that the ``baseline" and ``removal" models are not estimated separately; we estimate all parameters together, but when we generate predictions from the baseline model, 
we simply set the removal treatment effects to zero. We chose this over an alternative approach of fitting a model only to data from control plots and then making 
and evaluating predictions for the removal plots. The principal advantage of our approach is that the significance test (confidence intervals) for the removal effect 
is well-defined; fitting on one data set and testing on another would require developing non-standard null distributions to account for variation in both data sets. In addition, the alternative approach would return 
essentially the same results because adding data from the removal plots had negligible effects on other model parameters. Correlations between the parameters of 
a model fit only to control plot data and our full model (leaving aside the removal effects) were stronger than 0.999. 
Why don't we find strong interactions between removal treatments and the estimated competition coefficients? The answer is that because our treatments 
involved complete, not just partial, removal of target species, their effects can be captured with an intercept term, as illustrated in Fig. \ref{fig:PSSPresids}. 
As a result, if, for example, one of the grasses responds more positively to removal of \textit{A. tripartita} than our baseline model predicts, we cannot 
determine whether this reflects a poorly estimated competition coefficient, or some other effect of the removal treatment unrelated to competitor densities.

@@

As in our previous studies of this system \citep{adler_coexistence_2010,adler_forecasting_2012}, the kernel is constructed from the fitted survival ($S$), growth ($G$), and recruitment ($R$) models, 
\begin{equation}
k_j(v,u,\vec{n})=S_j(u,\vec{n})G_j(v,u,\vec{n})+R_j(v,u,\vec{n}) 
\label{eqn:IPM} 
\end{equation}
(see Appendix \ref{AppendixA} for a complete description). The IPM uses a mean-field approximation for
neighborhood competition that captures the essential features of spatially-explicit local interactions; 
in Appendix \ref{sec:kernelMethods} we explain how the spline competition kernels were used 
to compute neighborhood competition as a function of individual size for the IPM.

@@

During the period of our experiment, annual water-year (October-September) precipitation fell within the 5\% - 95\% quantiles of the 
long-term mean, although more years fell below the mean than above it (Fig. \ref{fig:climate}a in Appendix C). However, annual mean temperatures 
were consistently warm relative to the long-term mean, and two years were warmer than the 95\% quantile (Fig. \ref{fig:climate}b).

@@

Trends in cover by species and removal treatment were variable. For example, grass removal had little effect on the cover of \textit{A. tripartita} 
(Fig. \ref{fig:CoverTrends}), but shrub removal did appear to cause an increase in the cover of \textit{P. spicata}. 
However, when we tested the effects of removal treatments on year-to-year population growth rates ($\log(Cover_{t+1}/Cover_t)$) at the quadrat level,
 none of the removal treatment effects were significantly different from zero (Table \ref{table:coefficients}). For \textit{P. spicata}, 
the 95\% credible interval only just contained zero (95\% CI: -0.01, 0.38), providing marginal evidence for a positive effect of removal on population growth. 
Fitting with weighted least squares to address the high variance in residuals for plots with very low cover indicated even less evidence for a positive 
response of  \textit{P. spicata} to removal. However, these treatment comparisons do not directly address our questions about whether 
responses to removal were adequately predicted by models fit to observational data.

@@

Our vital rate models take into account the effects of local species composition and should therefore be able 
to explain the effect of removals on survival, growth and recruitment without any additional parameters. 
As in our previous studies using only the historical data, the neighborhood competition coefficients in the 
fitted models imply that interspecific competition is much weaker than intraspecific competition, in many
cases by 1-2 orders of magnitude (Table \ref{table:alphas}). The intraspecific interaction effects are large and have credible intervals that 
do not overlap zero, while the interspecific coefficients are small and the majority have credible intervals that do overlap zero.
If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of 
removals either larger or smaller than the effect predicted by a model based on data from the control plots. 
For the survival models, the 95\% credible intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; 
Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data 
from control plots. For the growth models, the 95\% credible intervals on the removal effects were positive for \textit{P. secunda}  and \textit{P. spicata},
 meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions 
 (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). \new{ At the recruitment stage, only \textit{A. tripartita} showed a significant response to removals, and it was negative (Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}).}

The positive removal effects on \textit{P. secunda}  and \textit{P. spicata} growth suggest that models based on observational data may underestimate competitive release. We conducted additional analyses to further explore this hypothesis.
First, \new{we asked whether increases in grass growth following removal were greatest in quadrats or neighborhoods where we removed the most \textit{A. tripartita}. We did this by taking residuals from the fitted growth model that included removal effects, and plotting the residuals for plants in removal quadrats against the pre-treatment, quadrat-level cover of \textit{A. tripartita}. No species showed the positive relationship that would be expected (Fig. \ref{fig:GrowthResidsQuad}).
In fact, \textit{Poa secunda} residuals were significantly lower in quadrats with high pre-treatment cover of \textit{A. tripartita}: individuals in high pre-treatment cover quadrats grew more slowly than predicted by the growth model with removal effects. We repeated this analysis at the individual plant level for grasses present in 2011. We calculated pre-treatment crowding from \textit{A. tripartita} for each of those individuals, then plotted their growth residuals against crowding for all years in which each plant survived. We found no significant relationships between growth residuals and pre-treatment crowding from \textit{A. tripartita} (Fig. \ref{fig:GrowthResidsIndivid}). }

Second, we evaluated the possibility that removal effects might be transient. We removed aboveground plant material, but roots of the removed plants remained in the ground potentially creating a flush of mineral nutrients as they decompose. We tested for a transient removal effect by adding a year-by-treatment interaction. For both  \textit{P. secunda}  and \textit{P. spicata}, we found no sign of a decrease in the removal effect over time (Tables \ref{table:POSEgrowth-trtYears} and \ref{table:PSSPgrowth-trtYears}). In fact, for  \textit{P. secunda}, the removal effects increased over time.

Third, responses to removals might be poorly predicted by models based on observational data if the removals reduce plant densities (crowding) below the range of historical variation, or if the removals result in combinations of neighborhood-scale densities that deviate from naturally occurring patterns of correlation. However, the distribution of crowding observed within control plots easily spans the crowding values observed in removal plots (Fig. \ref{fig:Wscatters}), and we failed to find evidence for strong interspecific covariances in crowding (Appendix \ref{AppendixA}).

Finally, we evaluated pre-existing differences in community composition in removal and control plots. In selecting removal plots, our criteria required candidate plots to include at least a certain amount of our target removal species. Because we adopted the historical plots as our controls, the same criteria were not applied (a control plot might not contain any \textit{A. tripartita}). To quantify resulting differences in pre-treatment (2011) composition, we compared neighborhood crowding at the individual plant level (the $W_{jm}$'s of the growth regressions) across treatments. \textit{A. tripartita} crowding and crowding by all subdominant shrub and grass species appeared greater in removal plots than in the controls, whereas crowding by \textit{H. comata} was lower in removals than controls, though it was absent from most \textit{P. spicata} neighborhoods in both treatments (Fig. \ref{fig:W-by-treatment}). These pre-existing differences in composition raise the possibility that plots selected for the removal treatments might differ from the control plots in some unobserved covariate that could influence plant composition and performance. \new{On the other hand, some of the control plots that currently contain zero \textit{A. tripartita} had high \textit{A. tripartita} in the 1900's, and vice versa. If abiotic variation, such as soil depth, were the primary control of variation in \textit{A. tripartita} cover, such large changes in cover over time would be unlikely. }

@@

Time series of predicted and observed cover, averaged over quadrats, show similar patterns. Predicted population size closely tracked observed population size (Fig. \ref{fig:IBM1step}) in most years for all species, but the difference between predictions from the baseline and removal models varied among species. For \textit{H. comata}, differences in predictions from the two models were virtually indistinguishable. For \textit{A. tripartita}, \textit{P. secunda} and especially \textit{P. spicata}, the model with removal effects consistently predicted higher population growth than the baseline model (Fig. \ref{fig:IBM1step}). Using the version of the model with maximum removal treatment effects, the patterns were qualitatively similar, with a somewhat greater contrast between predictions of the baseline and removal models for \textit{P. secunda} and \textit{P. spicata} (Fig. \ref{fig:IBM1step-maxCI}).

To quantify the long-term population-level consequences of permanent competitor removals, we used an IPM to project the long-term average cover of each species. The IPM predicts little effect of \textit{A. tripartita} removal on the cover of any grass species when we simulate the baseline model with no removal treatment coeffiicients (compare gray and blue boxes in Fig. \ref{fig:IPMresults}). In fact, projected average cover of all three grass species declines slightly following \textit{A. tripartita} removal. 
Using a model with removal treatment coefficients, the IPM projects that \textit{A. tripartita} removal will increase equilibrium cover, relative to the baseline scenario, by a factor of 1.4, 1.5 and 3.0 for \textit{H. comata}, \textit{Poa secunda} and \textit{P. spicata}, respectively (Fig. \ref{fig:IPMresults}). 
Running a version of the model with maximum removal treatment effects (removal coefficients set to the 95\% credible limit furthest from zero), resulted in a much larger projected increase in \textit{H. comata}, reflecting its small sample size and corresponding high uncertainty, but little change in the projections for \textit{Poa secunda} and \textit{P. spicata} (Fig. \ref{fig:IPMresults-maxCI}).

@@

The baseline model predicts more dramatic competitive release of \textit{A. tripartita} following grass removal: mean cover increases from 4\% to over11\%. 
Adding removal treatment effects dramatically alters the outcome, with \textit{A. tripartita} decreasing towards extinction. However, these outcomes both are driven by parameters that are estimated with very high uncertainty (Table~\ref{table:recruitment}). The baseline model predicts an increase following grass removal because of the relatively strong negative effects of \textit{Poa secunda} and \textit{H. comata} on \textit{A. tripartita} recruitment (\ref{table:recruitment}). The estimated effect of
\emph{Poa} is -0.50, but the standard deviation of the estimate is 0.28 and the 95\% credible interval stops just short of zero; the 95\% credible interval for the
\textit{H. comata} effect includes values less than half of the point-estimate. The grass removal treatment effect, which leads to predicted extinction, has a credible 
interval that broadly overlaps zero (Table~\ref{table:recruitment}). \textit{A. tripartita} recruitment parameters have exceptionally high uncertainty, presumably 
because \textit{A. tripartita} recruitment is episodic (see \emph{Discussion}), and this produces extreme uncertainty about its long-term 
response to grass removal.

@@

Our population model based on historical, observational data revealed weak interspecific interactions \citep{adler_coexistence_2010,chu_large_2015}. As a result, the model predicts very little competitive release among the four dominant species in a sagebrush steppe. 
Because the model was based solely on observational
data, we were suspicious of its ability to accurately describe the outcome of interspecific interactions, so we conducted a removal experiment to test it. 
Our approach represents a strong test of a surprising prediction for two reasons. First, the prediction was already published; the experiment 
gave us a very real opportunity to be wrong in public. Second, testing an observationally-based model with an experimental manipulation, 
not just additional observations, further increases the potential to confront the model with a novel situation outside the historical range of variation.
\new{We expected this study to provide clear evidence for the success or failure of our previous model's predictions. Unfortunately, our results do not provide that clarity. In the sections below, we first summarize the evidence that our experiment revealed greater competitive release than our baseline model predicted, then explain weaknesses in the experimental results that complicate the interpretation, and end by discussing general lessons for the study of interspecific interactions.}

@@

\new{We evaluated response to removals in three ways: vital rates, short-term population growth rates, and changes in long-term equilibrium cover. Two species, \textit{Poa secunda} and \textit{P. pseudoroegneria}, showed significant, positive responses to \textit{A. tripartita} removal at the level of vital rates. Individuals of these species grew 21\% and 24\% faster, respectively, in removal plots than controls, after accounting for variation in local neighborhood composition. In other words, our baseline model appeared to underestimate the competitive release of these species following \textit{A. tripartita} removal.}

\new{The elevated growth of \textit{Poa secunda} and \textit{P. pseudoroegneria} individuals translated into greater short-term population growth rates. For these species, the individual-based model with removal effects predicted higher cover than the baseline version of the model. The differences in predicted cover of roughly 20\% (Fig. \ref{fig:IBM1step}) are comparable to the differences in the individual growth rates, reflecting the fact that at these short time scales changes in individual size influence cover more than survival or recruitment.  For \textit{A. tripartita}, the magnitude of the difference between predictions from the baseline and removal effects models was similar, further emphasizing the short-term influence of growth compared to survival and recruitment, which responded negatively to removals in this species.   }

\new{Long-term projections using the integral projection model showed even larger responses to removal. For the three grasses, projected equilibrium cover in the absence of \textit{A. tripartita} is 1.4 to 3-fold higher for the removal effects model compared to the baseline model. While the absolute changes in basal cover are small, these relative changes are substantial, and seem to present compelling evidence that competitive release is stronger than predicted by our baseline model.  }

@@

\new{Although the growth regressions for \textit{Poa secunda} and \textit{P. pseudoroegneria} and the long-run projections for all three grass species suggest stronger competitive release than our baseline model predicted, a variety of other evidence makes it harder to confidently reject those baseline model predictions. Only two of the twelve vital rate tests indicated greater competitive release than predicted. Granted, small samples sizes contributed to greater uncertainty for \textit{A. tripartita} and \textit{H. comata}, but even for \textit{Poa secunda} and \textit{P. pseudoroegneria} the increases in growth were small relative to other sources of variation. For example, using quadrat-level cover data, we could not detect treatment differences (Table \ref{table:coefficients}).  }

\new{The strongest evidence that competitive release of grasses following \textit{A. tripartita} removal was greater than predicted comes from the long-term projections of the IPM. These projections effectively make removal treatment effects permanent; the removal effects on vital rates observed during the experiment are locked in for the entire 2500 time step simulation. Furthermore, these projections ignore parameter error because we wanted to highlight the population consequences of the removal effects. Given these strong assumptions, the projections should be viewed with some caution.  }

>>>>

\title{Weak interspecific interactions in a sagebrush steppe? Conflicting evidence from observations and experiments}

@@

Stable coexistence requires intraspecific limitations to be stronger than interspecific limitations. The greater the difference between intra- and interspecific limitations, the more stable the coexistence, and the weaker the competitive release any species should experience following removal of competitors. We conducted a removal experiment to test whether a previously estimated model, showing surprisingly weak interspecific competition for four dominant species in a sagebrush steppe, accurately predicts competitive release. Our treatments were 1) removal of all perennial grasses and 2) removal of the dominant shrub, \textit{Artemisia tripartita}. We regressed survival, growth and recruitment on the locations, sizes, and species identities of neighboring plants, along with an indicator variable for removal treatment. If our ``baseline" regression model, which accounts for local plant-plant interactions, accurately explains the observed responses to removals, then the removal coefficient should be non-significant. For survival, the removal coefficients were never significantly different from zero, and only \textit{A. tripartita} showed a (negative) response to removals at the recruitment stage. For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated interspecific competition. For all three grass species, population models based on the vital rate regressions that included removal effects projected 1.4 to 3-fold increases in equilibrium population size relative to the baseline model (no removal effects).
However, we found no evidence of higher response to removal in quadrats with higher pretreatment cover of \textit{A. tripartita}, or by plants experiencing higher pre-treatment crowding by \textit{A. tripartita}, 
raising questions about the mechanisms driving the positive response to removal.
While our results show the value of combining observations with a simple removal experiment, more tightly controlled experiments focused on underlying mechanisms may be required to conclusively validate or reject predictions from phenomenological models.

\textbf{\large{Keywords:}} Coexistence, competition, facilitation, integral projection model, population dynamics, removal experiment, sagebrush steppe.

@@

In spring 2011, in a large exclosure containing six of the historical permanent quadrats, we established an additional 16 quadrats for the removal treatments. New quadrats were selected to be similar to historical plots, but also needed to have enough of the focal species present so that removal experiments could be conducted. We evaluated potential locations for the new quadrats near the original quadrats and then used only those locations with at least 5\% canopy cover of \textit{Artemisia tripartita} and at least 3\% combined basal cover of the three focal perennial grasses. We rejected locations falling on hill slopes, areas with greater than 20\% bare rock, or with over 10\% cover of the woody shrubs \textit{Purshia tridentata} or \textit{Amelanchier utahensis}. We mapped the new quadrats in June, 2011 and then implemented the removal treatments in September, 2011. We randomly assigned 8 of the sixteen new quadrats to the sagebrush removal treatment, and the remaining 8 to the grass removal treatment. We cut sagebrush stems at ground level and applied a 5\% solution of Roundup Pro (Glyphosate, Monsanto) to the cut stems. In the grass removal plots, we painted the same herbicide solution on all perennial grasses. We removed plants 50 cm beyond the quadrat boundaries to minimize edge effects. The sagebrush removal was virtually 100\% effective; grass removals required additional herbicide applications in April and May of each year, with fewer grasses remaining each year. Appendix S1 Fig. \ref{fig:photos} shows photographs of example quadrats before and after treatment. We repeated mapped censuses of all plots each June from 2012-2016.

@@

Note that the ``baseline" and ``removal" models are not estimated separately; we estimate all parameters together, but when we generate predictions from the baseline model, 
we simply set the removal treatment effects to zero. We chose this over an alternative approach of fitting a model only to data from control plots and then making 
and evaluating predictions for the removal plots. The principal advantage of our approach is that the significance test (confidence intervals) for the removal effect 
is well-defined; fitting on one data set and testing on another would require developing non-standard null distributions to account for variation in both data sets. In addition, the alternative approach would return 
essentially the same results because adding data from the removal plots had negligible effects on other model parameters. Correlations between the parameters of 
a model fit only to control plot data and our full model (leaving aside the removal effects) were stronger than 0.999. 
Why don't we find strong interactions between removal treatments and the estimated competition coefficients? The answer is that because our treatments 
involved complete, not just partial, removal of target species, their effects can be captured with an intercept term, as illustrated in Appendix S1: Fig. \ref{fig:PSSPresids}. 
As a result, if, for example, one of the grasses responds more positively to removal of \textit{A. tripartita} than our baseline model predicts, we cannot 
determine whether this reflects a poorly estimated competition coefficient, or some other effect of the removal treatment unrelated to competitor densities.

@@

As in our previous studies of this system \citep{adler_coexistence_2010,adler_forecasting_2012}, the kernel is constructed from the fitted survival ($S$), growth ($G$), and recruitment ($R$) models, 
\begin{equation}
k_j(v,u,\vec{n})=S_j(u,\vec{n})G_j(v,u,\vec{n})+R_j(v,u,\vec{n}) 
\label{eqn:IPM} 
\end{equation}
(see Appendix S1 for a complete description). The IPM uses a mean-field approximation for
neighborhood competition that captures the essential features of spatially-explicit local interactions; 
in Appendix S1: \ref{sec:kernelMethods} we explain how the spline competition kernels were used 
to compute neighborhood competition as a function of individual size for the IPM.

@@

During the period of our experiment, annual water-year (October-September) precipitation fell within the 5\% - 95\% quantiles of the 
long-term mean, although more years fell below the mean than above it (Appendix S1: Fig. \ref{fig:climate}a). However, annual mean temperatures 
were consistently warm relative to the long-term mean, and two years were warmer than the 95\% quantile (Appendix S1: Fig. \ref{fig:climate}b).

@@

Trends in cover by species and removal treatment were variable. For example, grass removal had little effect on the cover of \textit{A. tripartita} 
(Fig. \ref{fig:CoverTrends}), but shrub removal did appear to cause an increase in the cover of \textit{P. spicata}. 
However, when we tested the effects of removal treatments on year-to-year population growth rates ($\log(Cover_{t+1}/Cover_t)$) at the quadrat level,
 none of the removal treatment effects were significantly different from zero (Appendix S1: Table \ref{table:coefficients}). For \textit{P. spicata}, 
the 95\% credible interval only just contained zero (95\% CI: -0.01, 0.38), providing marginal evidence for a positive effect of removal on population growth. 
Fitting with weighted least squares to address the high variance in residuals for plots with very low cover indicated even less evidence for a positive 
response of  \textit{P. spicata} to removal. However, these treatment comparisons do not directly address our questions about whether 
responses to removal were adequately predicted by models fit to observational data.

@@

Our vital rate models take into account the effects of local species composition and should therefore be able 
to explain the effect of removals on survival, growth and recruitment without any additional parameters. 
As in our previous studies using only the historical data, the neighborhood competition coefficients in the 
fitted models imply that interspecific competition is much weaker than intraspecific competition, in many
cases by 1-2 orders of magnitude (Appendix S1: Table \ref{table:alphas}). The intraspecific interaction effects are large and have credible intervals that 
do not overlap zero, while the interspecific coefficients are small and the majority have credible intervals that do overlap zero.
If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of 
removals either larger or smaller than the effect predicted by a model based on data from the control plots. 
For the survival models, the 95\% credible intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; 
Appendix S1: Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data 
from control plots. For the growth models, the 95\% credible intervals on the removal effects were positive for \textit{P. secunda}  and \textit{P. spicata},
 meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions 
 (Fig. \ref{fig:VitalRateTest}; Appendix S1: Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). At the recruitment stage, only \textit{A. tripartita} showed a significant response to removals, and it was negative (Fig. \ref{fig:VitalRateTest}; Appendix S1: Table \ref{table:recruitment}).

The positive removal effects on \textit{P. secunda}  and \textit{P. spicata} growth suggest that models based on observational data may underestimate competitive release. We conducted additional analyses to further explore this hypothesis.
First, we asked whether increases in grass growth following removal were greatest in quadrats or neighborhoods where we removed the most \textit{A. tripartita}. We did this by taking residuals from the fitted growth model that included removal effects, and plotting the residuals for plants in removal quadrats against the pre-treatment, quadrat-level cover of \textit{A. tripartita}. No species showed the positive relationship that would be expected (Fig. \ref{fig:GrowthResidsQuad}).
In fact, \textit{Poa secunda} residuals were significantly lower in quadrats with high pre-treatment cover of \textit{A. tripartita}: individuals in high pre-treatment cover quadrats grew more slowly than predicted by the growth model with removal effects. We repeated this analysis at the individual plant level for grasses present in 2011. We calculated pre-treatment crowding from \textit{A. tripartita} for each of those individuals, then plotted their growth residuals against crowding for all years in which each plant survived. We found no significant relationships between growth residuals and pre-treatment crowding from \textit{A. tripartita} (Appendix S1: Fig. \ref{fig:GrowthResidsIndivid}).

Second, we evaluated the possibility that removal effects might be transient. We removed aboveground plant material, but roots of the removed plants remained in the ground potentially creating a flush of mineral nutrients as they decompose. We tested for a transient removal effect by adding a year-by-treatment interaction. For both  \textit{P. secunda}  and \textit{P. spicata}, we found no sign of a decrease in the removal effect over time (Appendix S1: Tables \ref{table:POSEgrowth-trtYears} and \ref{table:PSSPgrowth-trtYears}). In fact, for  \textit{P. secunda}, the removal effects increased over time.

Third, responses to removals might be poorly predicted by models based on observational data if the removals reduce plant densities (crowding) below the range of historical variation, or if the removals result in combinations of neighborhood-scale densities that deviate from naturally occurring patterns of correlation. However, the distribution of crowding observed within control plots easily spans the crowding values observed in removal plots (Appendix S1: Fig. \ref{fig:Wscatters}), and we failed to find evidence for strong interspecific covariances in crowding (Appendix S1).

Finally, we evaluated pre-existing differences in community composition in removal and control plots. In selecting removal plots, our criteria required candidate plots to include at least a certain amount of our target removal species. Because we adopted the historical plots as our controls, the same criteria were not applied (a control plot might not contain any \textit{A. tripartita}). To quantify resulting differences in pre-treatment (2011) composition, we compared neighborhood crowding at the individual plant level (the $W_{jm}$'s of the growth regressions) across treatments. \textit{A. tripartita} crowding and crowding by all subdominant shrub and grass species appeared greater in removal plots than in the controls, whereas crowding by \textit{H. comata} was lower in removals than controls, though it was absent from most \textit{P. spicata} neighborhoods in both treatments (Appendix S1: Fig. \ref{fig:W-by-treatment}). These pre-existing differences in composition raise the possibility that plots selected for the removal treatments might differ from the control plots in some unobserved covariate that could influence plant composition and performance. On the other hand, some of the control plots that currently contain zero \textit{A. tripartita} had high \textit{A. tripartita} in the 1900's, and vice versa. If abiotic variation, such as soil depth, were the primary control of variation in \textit{A. tripartita} cover, such large changes in cover over time would be unlikely.

@@

Time series of predicted and observed cover, averaged over quadrats, show similar patterns. Predicted population size closely tracked observed population size (Fig. \ref{fig:IBM1step}) in most years for all species, but the difference between predictions from the baseline and removal models varied among species. For \textit{H. comata}, differences in predictions from the two models were virtually indistinguishable. For \textit{A. tripartita}, \textit{P. secunda} and especially \textit{P. spicata}, the model with removal effects consistently predicted higher population growth than the baseline model (Fig. \ref{fig:IBM1step}). Using the version of the model with maximum removal treatment effects, the patterns were qualitatively similar, with a somewhat greater contrast between predictions of the baseline and removal models for \textit{P. secunda} and \textit{P. spicata} (Appendix S1: Fig. \ref{fig:IBM1step-maxCI}).

To quantify the long-term population-level consequences of permanent competitor removals, we used an IPM to project the long-term average cover of each species. The IPM predicts little effect of \textit{A. tripartita} removal on the cover of any grass species when we simulate the baseline model with no removal treatment coeffiicients (compare gray and blue boxes in Fig. \ref{fig:IPMresults}). In fact, projected average cover of all three grass species declines slightly following \textit{A. tripartita} removal. 
Using a model with removal treatment coefficients, the IPM projects that \textit{A. tripartita} removal will increase equilibrium cover, relative to the baseline scenario, by a factor of 1.4, 1.5 and 3.0 for \textit{H. comata}, \textit{Poa secunda} and \textit{P. spicata}, respectively (Fig. \ref{fig:IPMresults}). 
Running a version of the model with maximum removal treatment effects (removal coefficients set to the 95\% credible limit furthest from zero), resulted in a much larger projected increase in \textit{H. comata}, reflecting its small sample size and corresponding high uncertainty, but little change in the projections for \textit{Poa secunda} and \textit{P. spicata} (Appendix S1: Fig. \ref{fig:IPMresults-maxCI}).

@@

The baseline model predicts more dramatic competitive release of \textit{A. tripartita} following grass removal: mean cover increases from 4\% to over11\%. 
Adding removal treatment effects dramatically alters the outcome, with \textit{A. tripartita} decreasing towards extinction. However, these outcomes both are driven by parameters that are estimated with very high uncertainty (Appendix S1: Table \ref{table:recruitment}). The baseline model predicts an increase following grass removal because of the relatively strong negative effects of \textit{Poa secunda} and \textit{H. comata} on \textit{A. tripartita} recruitment (\ref{table:recruitment}). The estimated effect of
\emph{Poa} is -0.50, but the standard deviation of the estimate is 0.28 and the 95\% credible interval stops just short of zero; the 95\% credible interval for the
\textit{H. comata} effect includes values less than half of the point-estimate. The grass removal treatment effect, which leads to predicted extinction, has a credible 
interval that broadly overlaps zero (Appendix S1: Table\ref{table:recruitment}). \textit{A. tripartita} recruitment parameters have exceptionally high uncertainty, presumably 
because \textit{A. tripartita} recruitment is episodic (see \emph{Discussion}), and this produces extreme uncertainty about its long-term 
response to grass removal.

@@

Our population model based on historical, observational data revealed weak interspecific interactions \citep{adler_coexistence_2010,chu_large_2015}. As a result, the model predicts very little competitive release among the four dominant species in a sagebrush steppe. 
Because the model was based solely on observational
data, we were suspicious of its ability to accurately describe the outcome of interspecific interactions, so we conducted a removal experiment to test it. 
Our approach represents a strong test of a surprising prediction for two reasons. First, the prediction was already published; the experiment 
gave us a very real opportunity to be wrong in public. Second, testing an observationally-based model with an experimental manipulation, 
not just additional observations, further increases the potential to confront the model with a novel situation outside the historical range of variation.
We expected this study to provide clear evidence for the success or failure of our previous model's predictions. Unfortunately, our results do not provide that clarity. In the sections below, we first summarize the evidence that our experiment revealed greater competitive release than our baseline model predicted, then explain weaknesses in the experimental results that complicate the interpretation, and end by discussing general lessons for the study of interspecific interactions.

@@

We evaluated response to removals in three ways: vital rates, short-term population growth rates, and changes in long-term equilibrium cover. Two species, \textit{Poa secunda} and \textit{P. pseudoroegneria}, showed significant, positive responses to \textit{A. tripartita} removal at the level of vital rates. Individuals of these species grew 21\% and 24\% faster, respectively, in removal plots than controls, after accounting for variation in local neighborhood composition. In other words, our baseline model appeared to underestimate the competitive release of these species following \textit{A. tripartita} removal.

The elevated growth of \textit{Poa secunda} and \textit{P. pseudoroegneria} individuals translated into greater short-term population growth rates. For these species, the individual-based model with removal effects predicted higher cover than the baseline version of the model. The differences in predicted cover of roughly 20\% (Fig. \ref{fig:IBM1step}) are comparable to the differences in the individual growth rates, reflecting the fact that at these short time scales changes in individual size influence cover more than survival or recruitment.  For \textit{A. tripartita}, the magnitude of the difference between predictions from the baseline and removal effects models was similar, further emphasizing the short-term influence of growth compared to survival and recruitment, which responded negatively to removals in this species.

Long-term projections using the integral projection model showed even larger responses to removal. For the three grasses, projected equilibrium cover in the absence of \textit{A. tripartita} is 1.4 to 3-fold higher for the removal effects model compared to the baseline model. While the absolute changes in basal cover are small, these relative changes are substantial, and seem to present compelling evidence that competitive release is stronger than predicted by our baseline model.

@@

Although the growth regressions for \textit{Poa secunda} and \textit{P. pseudoroegneria} and the long-run projections for all three grass species suggest stronger competitive release than our baseline model predicted, a variety of other evidence makes it harder to confidently reject those baseline model predictions. Only two of the twelve vital rate tests indicated greater competitive release than predicted. Granted, small samples sizes contributed to greater uncertainty for \textit{A. tripartita} and \textit{H. comata}, but even for \textit{Poa secunda} and \textit{P. pseudoroegneria} the increases in growth were small relative to other sources of variation. For example, using quadrat-level cover data, we could not detect treatment differences (Appendix S1: Table \ref{table:coefficients}).

The strongest evidence that competitive release of grasses following \textit{A. tripartita} removal was greater than predicted comes from the long-term projections of the IPM. These projections effectively make removal treatment effects permanent; the removal effects on vital rates observed during the experiment are locked in for the entire 2500 time step simulation. Furthermore, these projections ignore parameter error because we wanted to highlight the population consequences of the removal effects. Given these strong assumptions, the projections should be viewed with some caution.

For \textit{A. tripartita}, the long-term IPM projections actually show extinction following grass removal.
However, this result primarily reflects the influence of recruitment parameters that are highly uncertain (see \emph{Results}). This high uncertainty results from \textit{A. tripartita} recruitment being low and very episodic (roughly half of the total recruits in the 22 years of historical data were observed in 2 of those years, and only 28 \textit{A. tripartita} recruits have been observed since we resumed annual censuses in 2007). Because of this uncertainty, we put more weight on the direct observations and short-term projections showing little effect of grass removal on \textit{A. tripartita} populations.
